---
title: "PS3_Econ5043"
author: "Jordan Hoehne"
date: "3/7/2018"
output: word_document
---
```{r}
set.seed(123456)
setwd("/Users/jordan/desktop/")
library(psych)
library(randomForest)
library(caret)
#library(party)
```

```{r}
titanic <- read.csv("titanic.csv")
head(titanic)
```

```{r}
describe(titanic)
```

```{r}
summary(titanic)
```

```{r}
#Check for missing observations in the data 
sum(is.na(titanic))
```

```{r}
#Checking to see if there is any data missing in the columns we are working with 
sum(is.na(titanic$Pclass))
sum(is.na(titanic$Survived))
sum(is.na(titanic$Sex))
```
Since all of the data is there, we are good to proceed.

#Question 1
The statement, Pr[Survival | Pclass,Sex], can be expressed as:
Pr[PClass and Survived and Sex]/Pr[PClass and Sex].

```{r}
t_table<-table(Lived<-titanic$Survived, Gender<-titanic$Sex, Class<-titanic$Pclass)
t_joint_dist<-prop.table(t_table)
t_joint_dist
```

```{r}
t_margin_dist<-addmargins(prop.table(t_joint_dist))
t_margin_dist
```

```{r}
#par(mfrow=c(1,1))
#simple_survival_ctree <- ctree(Survived ~ Pclass + Sex, data=titanic, controls=ctree_control(testtype="Bonferroni"))
#plot(simple_survival_ctree)
```

#Question2
#Distribution = (9H, 3T)

Let’s assume the following priors (f(θ)): 
p(θ = .15) = .3 
p(θ = .25) = .2
p(θ = .5) = .3
p(θ = .75) = .2

```{r}
#setting variables to work with later 
p_theta.15 <- .15
p_theta.25 <- .25
p_theta.5 <- .5
p_theta.75 <- .75
```

Calculating the likelihood (the probability of coming up with a H is θ, then that of T is 1 − θ):
(a) What is the general formula for f(x|θ)?
  The general formula is: p^n*(1-p)^N-n, for N observations 
(b) Now calculate f(x|0.15),f(x|0.25),f(x|0.5),f(x|0.75)
```{r}
p.t.15<-p_theta.15^9*(1-p_theta.15)^3
p.t.25<-p_theta.25^9*(1-p_theta.25)^3
p.t.50<-p_theta.5^9*(1-p_theta.5)^3
p.t.75<-p_theta.75^9*(1-p_theta.75)^3
p.t.15
p.t.25
p.t.50
p.t.75
```

Calculating f(x) using the law of total probability
```{r}
f_x1<-p.t.15*.3
f_x2<-p.t.25*.25
f_x3<-p.t.50*.3
f_x4<-p.t.75*.2
f_x<-sum(f_x1 + f_x2+f_x3+f_x4)
list(f_x1, f_x2, f_x3, f_x4, f_x)
```

Calculating the posterior distribution:
(a) f(θ1 = .15|x) = f(x|θ1)·f(θ1) / f (x)
(b) f(θ2 = .25|x) = f(x|θ2)·f(θ2) / f (x)
(c) f(θ3 = .50|x) = f(x|θ3)·f(θ3) / f (x)
(d) f(θ4 = .75|x) = f(x|θ4)·f(θ4) / f(x)

```{r}
#a
a<-f_x1/f_x
#b
b<-f_x2/f_x
#c
c<-f_x3/f_x
#d
d<-f_x4/f_x
#Each Probability 
list(a,b,c,d)
#Total Probability 
Total_Probability<-sum(a,b,c,d)
Total_Probability

```


```{r}
options(digits=3)
Prior_Posterior<- matrix(c(.3, .2, .3, .2, a, b, c, d), ncol=4, byrow=TRUE)
colnames(Prior_Posterior) <- c("θ1", "θ2", "θ3", "θ4")
row.names(Prior_Posterior) <- c("Prior", "Posterior")
Before_After<-as.table(Prior_Posterior)
Before_After
```
Given 12 flips of a coin, we believed we would the results for .15, .25, .5, and .75 of the results to return heads as .3, .2, .3, and .2. The resulting probabilities were: 0.000023, 0.001305, 0.237575, and 0.761097 respectively. It is unlikely to see only .15 or .25 of the flips lead to 9 Heads and 3 Tails and we are more likely to see .5 of the coin flips to return in 9 Heads and 3 Tails and even more likely for .75 of the coin flips to result in 9 Heads and 3 Tails. The differnce could be in a small amount of data collected for our priors, which would explain the difference. 

#Question 3
#Interpreting the Medical Test (I)
A Dr. says that 90 percent of the people (!) who have smallpox have the symptoms you have. Since smallpox is often fatal, your first inclination may be to panic. But your doctor would tell you that the probability you have smallpox is only 1.1 percent (or 0.011), so you do not have to panic.

###Your doctor’s statements seem contradictory to each other. How to reconcile it? Is your doctoring just comforting you? 
The Dr.'s statements are not contradicting one another. He has made 2 statements that both pertain to your medical conditions.
It's very possible that you have fallen into the 1.1% of individuals that have smallpox. Given that you 90% of the symptoms,
you should be concerned for you health as you likely have smallpox. Since the Dr. did not state that other illnesses share
these symptoms, I would advise you meet with your Dr again and ask for further testing.

###Or, is there any scientific reason to explain the seemingly contradictory statements that you doctor made?
There are possible explanations for his diagnosis. He could be a terrible Dr. or there is a slight possibility that 
you do not have small pox. Even though you have share 90% of the symptoms, you could not have smallpox. Alternatively,
there may have been other cases where someone only shared 50% of the symptoms or perhaps showed no symptoms  at all 
and still had smallpox. 

#Question 4
#Interpreting the Medical Test (II)
```{r}
#Establish Probabilities 
#Identical Twins are always the same sex
#Fraternal Twins have a P(.5) of being the same sex 
twins<-matrix(c(1/3,.5*2/3,0,.5*2/3), ncol=2,byrow=TRUE)
colnames(twins) <- c("Identical Twins", "Fraternal Twins")
row.names(twins) <- c("Same_Sex", "Different_Sex")
twins<-addmargins(as.table(twins))
twins
```
Thus, the physician should inform the individual that she has a 1/3 chance that her twins will be identical and have the same sex 

#Question 5 
#Bayes’ Rule and Hiring Decisions
```{r}
All_Employees<-1
Qualified_Employees<-.80
Unqualified_Employees<-.20
Qualified_Interviewed_Well<-.90
Qualified_Interviewed_Poorly<-.10
Unqualified_Interviewed_Well<-.20
Unqualified_Interviewed_Poorly<-.80
```

Now, based on this data, what is the probability of a job candidate who interviewed well being a qualified worker?
```{r}
interview<-matrix(c(.8*.9,.2*.2,.8*.1,.2*.8), ncol=2,byrow=TRUE)
colnames(interview) <- c("Qualified Worker", "Unqualified Worker")
row.names(interview) <- c("Interviewed Well", "Interviewed Poorly")
interview<-addmargins(as.table(interview))
interview
```
With the given table, we can find out the odds that someone that interviewed well being qualified as Qualified and Interviewed Well / Interviewed Well 
```{r}
Prob_Interviewed_Well_and_Qualified <- .72/.76
Prob_Interviewed_Well_and_Qualified
```

```{r}

```

