---
title: "HW4_Redo"
author: "Jordan Hoehne"
date: "5/9/2018"
output: word_document
---
---
title: "HW4_Hoehne.R"
author: "Jordan Hoehne"
date: "4/6/2018"
output: word_document
---

#Load Packages
```{r}
library(MASS)
library(psych)
library(ggplot2)
```

#Simulate Data & Set Seed 
```{r, results="hide"}

#set seed for reproducability 
set.seed(123456)

#sample data 
y <- sample(c(1:5),1000, replace= TRUE, prob=c(.25,.25,.2,.2,.1)) 
x <- sample(c(0:1),1000, replace= TRUE, prob=c(.5,.5))
z1 <- sample(c(0:1),1000, replace= TRUE, prob=c(.25,.75))
z2 <- sample(c(0:1),1000, replace= TRUE, prob=c(.3,.6))

#create data frame for sampled variables 
data <- as.data.frame(cbind(y,x,z1,z2))
attach(data)
```

#Manually select a subsample and calculate the conditional distribution.
```{r}
#subset data
data.xo<-subset(data, x==0, select=y)

#joint probability for data.xo
prop.table(table(data.xo))
```

```{r}
#subset the data
data.x1<-subset(data, x==1, select=y)

#joint probability for data.x1
prop.table(table(data.x1))
```

#Generate dummy variables for each category of Y .
```{r}
#create new column in data 
#use "factor(var) +0" to create factor variable for analysis 
data$dummies <- model.matrix(~factor(y) + 0, data = data)
```

#use the aggregate() function to calculate the conditional distribution.
```{r}
#for x = 0, predict y
#for x = 1, predict y
aggregate(data$dummies, by=list(x), FUN=mean)
```

#Prediction for Y when X = 1?
The prediction for Y when x=1, my prediction of y is 1, given y1 has the largest probability of the y values.

#Use whichever approach that you like to calculate the conditional distribution Pr[Y |X, Z1, Z2].
```{r}
#every posible makeup of variables 
aggregate(data$dummies, by=list(x,z1, z2), FUN=mean)
```

#using your results in 2), what is your prediction when X = 1, Z1 = 1, Z2 = 0?
My prediction for Y when x = 1, z1 = 1, z2 = 0 is y3 as the probability of y3 is 0.2857143 and has 
the largest probability for the y value.

```{r}
#load airbnb data 
airbnb<-read.csv("airbnb.csv")
names(airbnb)
```

#If a customer requests a room with a check-in gap of 11 days, is it likely for this person to get his request accepted? Why?
```{r}
#check all listing that were requested with an 11 day gap
data.bnb<-subset(airbnb, checkin.gap==11)

#view data 
data.bnb
```
Given that 3 of the 4 requests were denied, there is a 75% chance that the request will be denied.

#Suppose that a customer submitted his request. Coincidentally, every host in your dataset has a check-in gap of 2 days for this particular request. 
#Which host(s) should you recommend to this customer in order to maximize the chance that this recommended host will accept this request?
```{r}
data.bnb2<-subset(airbnb, checkin.gap==2)
data.bnb2
```
Given that the Miller and Smith residence accepted requests with a 2 day check in gap, where other residences did not, I would recomend that guests request to book with the Miller and Smith residence as requests with a 2 day check in gap were accepted. 

#Simulate Data 
```{r}
#set seed again for reproducability 
set.seed(123456)

#set normal distribution with rnorm
yy <- rnorm(1000)

#set distribution of zz with values 1 through 3 
zz <- sample(1:3,1000, replace = TRUE)

#create new table from the set values 
data2 <- as.data.frame(cbind(yy, zz))
```

#Set up data to be analyzed
```{r}
#set dummy variable using factor(var) + 0 
data2$dummies <- model.matrix(~factor(y) + 0, data = data2)
```

#Calculate the conditional mean function with aggregate() command. And write it in one equation.
```{r}
#conditional means for data
aggregate(data2$dummies, by=list(yy, zz), FUN=mean)
```

#Similarly calculate the conditional median with two approaches.
#Subset approach (manually)
```{r, results="hide"}
#using subset instead of aggregate, provides propbability of each observation
data2.1<-subset(data2, zz==1, select=yy, FUN=median)

#view probability of each observation
prop.table(table(data2.1))
```

```{r, results="hide"}
#using subset instead of aggregate, provides propbability of each observation
data2.2<-subset(data2, zz==2, select=yy, FUN=median)

#view probability of each observation
prop.table(table(data2.2))
```

```{r, results="hide"}
#using subset instead of aggregate, provides propbability of each observation
data2.3<-subset(data2, zz==3, select=yy, FUN=median)

#view probability of each observation
prop.table(table(data2.3))
```

#Use aggregate(). You need to learn how to specify a different function in this case.
```{r}
#aggregate is an automatic way of determinig conditional mean
aggregate(data2$dummies, by=list(yy, zz), FUN=median)
```

#Suppose that this is some type of financial data for which I care about the lower-tail risks, and I would like to predict such value. 
#z represents different markets. 
#Suppose that I would like know about 5th percentile of my returns to this stock in the market z = 1. Calculate this value.

#summary stats for different markets
As you can see, the markets are not so different.
```{r}
df<-data.frame(a=as.data.frame(summary(data2.1)),b=as.data.frame(summary(data2.2)),c=as.data.frame(summary(data2.3)))
names(df) <- c("a.Var1", "a.Var2", "z==1", "b.Var1", "b.Var2", "z==2", "c.Var1", "c.Var2", "z==3")
keep <- c("z==1","z==2","z==3")
df[keep]
```

Here we can view the different markets which z represents 
```{r}
ggplot(data=data2, aes(x=yy, y=zz)) + geom_point(aes(color = factor(zz))) + ggtitle("Different Markets") + theme_minimal() + scale_color_discrete(name="zz", labels=c("zz==1", "zz==2", "zz==3"))
```


```{r}
#Using data subset with z=1 set up earlier
lowrisk<-quantile(data2.1, 0.05, type=1, na.rm=TRUE)
lowrisk
```

The resulting calculation of the 5th percentile, thus implies that we expect to lose no more than -1.628805 only 5% of the time.

By plotting the data with a histogram in ggplot, we can get a better picture of what's going on. 
```{r}
#using subset of data for z==1 
ggplot(data=data2.1, aes(x=yy)) + geom_histogram() +theme_minimal()
```
