---
title: "HW1"
author: "Jordan Hoehne"
date: "2/5/2018"
output:
  word_document: default
  pdf_document: default
---
###Please note I worked with Alex Skipper to help think through the logic on some of the questions 

#1 Load prework configurations 
```{r}
#set seed for reproducability
set.seed(123456)

#vcd needed for mosaic plot
library(vcd)

#load in data 
titanic<-read.csv("titanic.csv")

```
#making prop table to see joint probabilities
```{r}
#setting new columns
survived<-(titanic$Survived)
pclass<-(titanic$Pclass)

#set new table for analysis 
titanic_table<-table(survived,pclass)

#data<-as.data.frame(cbind(survived,pclass))

#set prop table 
prop_table<-prop.table(titanic_table)

#view prop table 
prop_table

```

#Add margins to see marginal probabilities for columns and rows
```{r}
addmargins(prop_table)
```

#mosaic plot of table 
```{r}
mosaic(titanic_table, shade=TRUE, legend=TRUE)
```


#xtabs used to create contingency table 
```{r}
titanic_xtabs<-xtabs(~survived+pclass)
titanic_xtabs
addmargins(titanic_xtabs)
```

#4 probability of pclass 1 and 2 
```{r}
spclass1and2<-(136+87)
spclass1and2/891
```

#5
```{r}
Expected<-table(titanic$Survived, titanic$Pclass)

#view expected 
Expected

#add margins to expected value 
addmargins(Expected)

#test significance of expected 
chisq.test(Expected)

#manual comparison
Pclass1<-(216/891*549/891)*891
Pclass2<-(184/891*549/891)*891
Pclass3<-(491/891*549/891)*891

P11<-(216/891*342/891)*891
P12<-(184/891*342/891)*891
P13<-(491/891*342/891)*891

#percentage
Pclass1/549
P11/342
Pclass2/549
P12/342
Pclass3/549
P13/342

#p-value is low, thus reject the null hypothesis. Essentially, the results  show that the classes are not independent. 

#you could use mosaic to help view independence, you could also test the squared errors through angular testing to see indepence 
```

#create table for probabilities 
```{r}
library(data.table)
row_names<-c("70","75","80")
col_names<-c("30","32","34")
Q2_matrix<-matrix(c(.1,.2,0,.1,.2,.2,0,0,.1),nrow=3,ncol=3,dimnames=list(row_names,col_names))
Q2_matrix
#Viewing the margins 
addmargins(Q2_matrix)
chisq.test(Q2_matrix)
```
#Q2.2

#probability is 0.6 on the return of stock A: x=<75

#Q2.3

#Unable to make a prediction on the prob of her portfolio that it is actually worth 13,000. 



